---
title: "SA2-DELA ROSA, R"
author: "Roland Dela Rosa"
date: "`r Sys.Date()`"
output: md_document
subtitle: "**Github Link: 
https://github.com/rddelarosa/APM1111/blob/main/SA2/DELA%20ROSA%2C%20R-SA2.Rmd**"
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(ggplot2)
library(dplyr)
library(car) # For Levene's test and Anova()
```

```{r}

# Load the Data
df <- read.csv("Alzheimers Mice Data.csv")

df$AD_Status <- as.factor(df$AD_Status)
df$Treatment <- as.factor(df$Treatment)

# Check the structure to ensure conversion worked
str(df)
```

```{r}

# Descriptive Statistics
summary_stats <- df %>%
  group_by(AD_Status, Treatment) %>%
summarise(
  Mean_Memory = mean(Memory),
  SD_Memory = sd(Memory),
  Mean_Training = mean(Training),
  Count = n(),
  .groups = "drop"
)
print(summary_stats)
```

```{r}
#  Visualization (Boxplot)
ggplot(df, aes(x = Treatment, y = Memory, fill = AD_Status)) +
  geom_boxplot() +
  labs(title = "Effect of Treatment and AD Status on Memory",
       x = "Treatment Type",
       y = "Memory Score") +
  theme_minimal()
```

```{r}

# Homogeneity of Variance (Levene's Test)
leveneTest(Memory ~ AD_Status * Treatment, data = df)

options(contrasts = c("contr.sum", "contr.poly"))
anova_model <- aov(Memory ~ AD_Status * Treatment, data = df)
shapiro.test(residuals(anova_model))

qqnorm(residuals(anova_model))
qqline(residuals(anova_model))

```

```{r}
Anova(anova_model, type = "III")
TukeyHSD(anova_model, which = "Treatment")

```

A two-way analysis of variance (ANOVA) was conducted to examine the effects of Alzheimer’s disease status (AD vs. non-AD) and treatment condition (four levels) on memory performance.

There was a significant main effect of AD status on memory scores,
F(1, 32) = 75.31, p < .001, indicating that mice with Alzheimer’s disease differed significantly in memory performance compared to non-AD mice.

The main effect of treatment was not significant,
F(3, 32) = 1.92, p = .146, suggesting that memory scores did not differ significantly across the four treatment conditions.

The interaction between AD status and treatment was also not significant,
F(3, 32) = 1.15, p = .344, indicating that the effect of treatment on memory performance did not depend on Alzheimer’s disease status.

Because the main effect of treatment was not significant, post hoc Tukey HSD tests revealed no significant pairwise differences between treatment groups.

